Exploatarea Software și Mai Departe: Dezvăluirea Vectorilor de Atac și Contra-Măsurilor în Calculul Modern / Software Exploitation and Beyond: Unveiling Attack Vectors and Countermeasures in Modern Computing

### 2023-06-24 paper.tex
wowieww

@@

\section{Detection techniques}%

\section{Mitigation Techniques}%

\section{Case studies}%

>>>>

If we do not possess the required information to execute our exploit, we can use
another one (another link in the exploit chain) to gather more and more. This
can be interpreted as going further into unintended states in the IFSM (intended
finite state machine) described here\ref{add to weird machines}.

@@

Driven by the relentless demand for heightened performance, a detailed scrutiny
of processor operations and employed microarchitectural designs has become
imperative. This necessity is underscored by the unearthing of vulnerabilities
such as those referred to in \cite{spectre, meltdown}. These exploits
ingeniously capitalize on the side effects triggered by the processor during
memory access and speculative execution. This development highlights the
criticality of an in-depth understanding of hardware function. Even before
the findings mentioned earlier, researchers have been able to cross the
virtualization boundary and gain information form other VM's running in the same
machine. The \cite{flush reload} reveals an attack that can recover around \~95\%
of the bits of a GnuPG secret key.

\section{Detection and Mitigation Techniques}%
In the realm of cybersecurity, the price tag associated with the implementation
of mitigation strategies such as stack canaries and heuristic checks is
frequently overlooked. These countermeasures, while crucial for system
protection, impose an overhead on system performance, potentially influencing
the overall efficiency of the software. As part of our investigation, this paper
will delve into an in-depth analysis of this often-neglected cost associated
with such defensive measures. Our objective is to facilitate a more balanced
perspective, considering not only the protective benefits but also the incurred
costs of these mitigation strategies.

\subsubsection{Stack canaries}
Defending against stack buffer overflows has been attempted all the way since 1997
with StackGuard \cite{stackguardpaper}, but it wasn't until 2006 with the release
of GCC 4.1 was this available for the entire Linux community.

In modern compilers this mitigation places a sacrificial random value right before the
return address. This mitigation combined with a new variable alignment technique that
places all buffers above all other function local variables has proven to be an
effective mitigation with a minimal performance impact.

\begin{lstlisting}[
      ,language=C]
#include <string.h>
extern void dont_optimize_str(char* ptr);
extern void dont_optimize_int(int* ptr);

int buffer_first(char* source) {
    char dest[0x100];
    int len;
    int foo;
    int bar;

    strcpy(dest, source);
    len = strlen(dest);
    foo = len * 2;
    bar = len - 3;

    dont_optimize_str(dest);
    dont_optimize_int(&len);
    dont_optimize_int(&foo);
    dont_optimize_int(&bar);
}

int buffer_last(char* source) {
    int len;
    int foo;
    int bar;
    char dest[0x100];

    strcpy(dest, source);
    len = strlen(dest);
    foo = len * 2;
    bar = len - 3;

    dont_optimize_str(dest);
    dont_optimize_int(&len);
    dont_optimize_int(&foo);
    dont_optimize_int(&bar);
}
\end{lstlisting}%
\label{code:example_func}

On GCC x86-64 version 13.1, compiling the code\ref{thatcode} results in both functions having the buffer
placed closest to the stack canary.

\begin{lstlisting}[caption={Shellcode disassembly}
      ,label={lst:bof}
      ,language={[x86masm]Assembler}]
        mov     rdx, QWORD PTR [rbp-296]
        lea     rax, [rbp-272]
        mov     rsi, rdx
        mov     rdi, rax
        call    strcpy
\end{lstlisting}

\begin{lstlisting}[caption={Shellcode disassembly}
      ,label={lst:bof}
      ,language={[x86masm]Assembler}]
        mov     rax, QWORD PTR fs:40
        mov     QWORD PTR [rbp-8], rax
\end{lstlisting}

\begin{lstlisting}[caption={Shellcode disassembly}
      ,label={lst:bof}
      ,language={[x86masm]Assembler}]
        mov     rdx, QWORD PTR [rbp-8]
        sub     rdx, QWORD PTR fs:40
        je      .L2
        call    __stack_chk_fail
\end{lstlisting}

\subsection{Performance cost}
\begin{table}[ht]
\centering
\begin{tabular}{|l|r|r|r|r|r|r|r|}
\hline
Compiler flags & Mean & $\delta$ & Min & Max & Speedup \\
\hline
 Optimize &  47.8 ms & 2.8 ms & 43.6 ms & 62.3 ms & 1x \\
 No stack protector&  46.6 ms & 1.9 ms  & 43.2 ms  &  59.1 ms & 1.025x \\
 Ubsan &  102.8 ms & 3.4 ms & 95.6 ms  & 116.5 ms & 0.460x \\
\hline
\end{tabular}
\caption{GCC results}
\label{table:your_label}
\end{table}

\begin{table}[ht]
\centering
\begin{tabular}{|l|r|r|r|r|r|r|r|}
\hline
Compiler flags & Mean & $\delta$ & Min & Max & Speedup \\
\hline
 Optimize &  47.6 ms & 2.1 ms & 44.0 ms & 68.8 ms & 1x \\
 Ubsan &  119.4 ms & 5.3 ms &  112.6 ms & 150.8 ms & 0.398x \\
 Ubsan minimal & 72.0 ms & 3.6 ms &  66.0 ms& 89.2 ms & 0.661x \\
\hline
\end{tabular}
\caption{Clang results}
\label{table:your_label}
\end{table}

Mention inability to test shadow stack

\subsection{Shadow stacks}
Advancements in secure software systems have led to the development of shadow
stacks, which represent a significant leap forward from the earlier protection
measures like stack canaries. Shadow stacks offer a more robust mechanism for
guarding against stack buffer overflows by completely decoupling control-flow
information from user data. They maintain a separate, concealed stack dedicated
solely to storing return addresses, making it exceedingly challenging for
attackers to alter the flow of execution. This strategic bifurcation of data
fundamentally strengthens the protection against control-flow hijacking attacks.
It ensures that even if an attacker successfully tampers with data on the main
stack, any discrepancies with the untouched shadow stack will immediately be
flagged. Despite the operational complexities and slight runtime overheads they
may introduce, shadow stacks serve as a remarkably effective mitigation
strategy, enhancing the integrity of the control-flow beyond what is achievable
with stack canaries alone.

The security and performance implications of shadow stacks are discussed in depth
in \ref{shadow stacks}

Compile TCC here with shadow steck and show perf diff.

\subsection{Sanitizers}
With the increasing amount of memory errors happening in unsafe system's programming
languages such as C and C++, methods for detecting runtime errors without requiring
code changes have been developed. Even though they are rarely used in production
environments they have seen massive success when used in conjunction with fuzzers
\ref{vezi niste paperuri de fuzzere} to detect severe vulnerabilities in really
large code bases.

\subsubsection{UBSAN}
UBSAN (Undefined Behavior Sanitizer) injects additional checks during the compilation of
the program in order to detect Undefined Behaviour\footnote{explain this}. Here is a
non-exhaustive list of the features and checks UBSAN does:

\begin{itemize}
  \item Array subscript out of bounds, where the bounds can be statically determined
  \item Bitwise shifts that are out of bounds for their data type
  \item Dereferencing misaligned or null pointers
  \item Signed integer overflow
  \item Conversion to, from, or between floating-point types which would overflow the destination
\end{itemize}

Give example with shellcode example not vorking

Compile TCC with and without ubsan(pls god work)

\subsubsection{Seccomp}
Seccomp (or Secure Computing) was introduced in the 2.6.12 version of the
Linux kernel. At that time it provided a \emph{strict} execution mode for
processes that allowed a very limited set of system calls. The intention
behind this mode was to allow untrusted code to run on systems. This
functionality was greatly enhanced with the introduction of seccomp-bpf
(Berkeley Packet Filter) in version 3.5 of the kernel. This new mode of
operation allowed programmers to filter the systemcalls a process makes
using small programs written in a DSL. This functionality sees proeminent
use in projects where the security of the host system is paramount, such as
Docker, Chrome or Firefox.

Here\ref{codubun} is an example of the granular control seccomp can achieve.
Note that once a process install a seccomp filter it is impossible to remove
it without exploiting the kernel, while also being inherited by all child
processes of the filtered process.

show example of mitigations adding new vulnerabilities (jit attack)
\section{Economic impact}%
Now that the performance impact of the mitigations has been established, we
will take a look at the economic/financial impact of the exploits these
mitigations are trying to prevent.

The domain of blockchain has emerged as one of the most publicised arenas for
software exploitation in recent years. As a technology built upon the principles
of decentralization and cryptography, blockchain applications, especially
cryptocurrencies, have attracted significant attention from both ethical and
malicious actors. The incentive for exploiting these systems is often financial
gain, given the tangible real-world value of many digital assets. Smart contract
vulnerabilities, insecure wallets, and compromised blockchain platforms have led
to numerous notable breaches. This highlights the critical importance of
rigorous security practices in the design, implementation, and operation of
blockchain-based systems. Multiple publications estimated that in 2022 alone
4 billion dollars have were stolen.

Ransomware has also emerged as a significant threat in the software exploitation
arena, causing extensive harm primarily through service disruptions. Beyond
direct financial loss from ransoms, these attacks result in considerable
operational downtime and ensuing economic damage. Consequently, the real impact
of ransomware extends beyond the immediate ransom, emphasizing the necessity of
comprehensive security measures and incident response strategies.
